# JOUR405: Statistics for Journalists
## Midterm Exam - Spring 2025

Name: William Deffaa

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences. Start by loading the tidyverse and any other libraries you think you might need.

```{r}
library(tidyverse)
```


## Part 1: Restaurant Health Inspections (15 points)

You want to understand how restaurants in Montgomery County are performing on health inspections. The first dataset contains restaurant health inspection scores for restaurants in Montgomery County. The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv` and complete these tasks:

```{r}
health_score <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```


### Tasks:
1. Calculate the mean and standard deviation of compliance scores (5 points)
2. Create a histogram of the compliance scores with a vertical line showing the mean (5 points)
3. Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? (5 points).

The standard deviation an the histogram tell us that the distribution of higher compliance scores skews more than scores that are low, with very few restaurants having compliance scores below 90 and many having compliance scores of 100. The most news worthy thing about this dristribution of resturants in MOCO and their compliance scores is that many of these resturants have very high scores, with only a select few having lowerish scores for compliance. 


```{r}
health_score %>% summarize(sd_compliance = sd(compliance_score), mean = mean(compliance_score))
  health_score %>% 
    ggplot(aes(x = compliance_score)) +
    geom_histogram(binwidth = 5.0, fill = "gray", color = "black") +
    geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed") +
    labs(title = "Distribution of Compliance Scores")
  
```

## Part 2: High School Athletics (25 points)

You are reporting a story about high school sports participation in Maryland and want to see if there are differences between boys and girls. The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and sex. Load the data from: `https://raw.githubusercontent.com/example/md_hs_sports_2024.csv` and complete these tasks:

```{r}
hs_sports <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv") 
```


### Tasks:
1. Calculate the correlation between boys' and girls' participation (5 points)
2. Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. (5 points)
3. Create a scatterplot showing this relationship, adding a line of best fit (5 points)
4. In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

```{r}
hs_sports %>% summarize(correlation = cor(boys, girls, method = "pearson"))
hs_sports <- hs_sports %>% mutate(total = boys + girls)
hs_sports <- hs_sports %>% mutate(girls_pct = girls/total)
  hs_sports %>%
    ggplot() +
    geom_point(aes(x = girls_pct, y = total)) +
    geom_smooth(aes(x = girls_pct, y = total), method = "lm") +
    geom_text(aes(x = girls_pct, y = total, label = district), size = 2)
```
The scatterplot reveal that there is not much equity in schools that don't have a lot of students, with the main outlier being Baltimore County Public Schools which has 70% girls in High School Sports. School districts that are above the line typically have more total students but have low equity in high school sports.


## Part 3: Public Transit Ridership (20 points)

You are investigating public transit ridership in the Washington, D.C. area and want to understand the patterns of daily bus and rail ridership. The third dataset contains daily bus and rail ridership totals from WMATA for the past year. Load the data from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv and do the following:

```{r}
WMTA <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/wmata_daily.csv")
```


### Tasks:
1. Calculate the average bus and rail ridership and standard deviation using summarize() (5 points)
2. Using the process you used in class, take a random sample daily ridership numbers and calculate the sample means and deviations for bus and rail. The number in the sample is up to you, but explain why you chose what you did. Compare this to the stats you generated in step 1. (5 points)
3. Using group_by() and summarize(), calculate the means for bus and rail ridership for each weekday. Describe the overall pattern of ridership for bus and rail - which days stand out and why? Are there differences between bus and rail in the standard deviation values? (10 points)


```{r}
WMTA %>% summarize(sd_bus = sd(bus), sd_rail = sd(rail))
  sample_WMTA <- WMTA %>% sample_n(100)
  sample_WMTA %>% summarize(sd_bus_sample = sd(bus), sd_rail_sample = sd(rail))
  sample_WMTA %>% summarize(mean_bus_sample = mean(bus), mean_rail_sample = mean(rail))
WMTA %>% group_by(weekday) %>% 
    summarize(wk_bus = mean(bus), wk_rail = mean(rail))
    
```
The standard deviation for rail is overall higher than for bus, but they are still relatively close in standard deviation. In terms of mean and what weekdays standout the most, Tuesday, Wednesday and Thursday are the highest in terms of mean riders for the week, with Sunday and Saturday being the lowest. I am suprised that Monday and Friday are not at the top for both rail and bus. 


## Part 4: Maryland Car Theft Rates (20 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

```{r}
MD_theft <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```


### Tasks:
1. Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate (5 points)
2. Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median, and what percentage of all car thefts occur in those counties? (5 points)
3. Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? (10 points)


```{r}
MD_theft <- MD_theft %>% mutate(per_capita_rate = 2023/population)
  MD_theft %>% summarize(median_PCR = median(per_capita_rate))
```

## Part 5: Data Analysis Scenario (20 points)

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences (no code!) explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)

I would measure the reponse times by when the call was made to when service would arrive, I would group these into police, firefighters, and EMTs. I would measure the median response time of the year that is being discussed and compare it to years prior to see if the response time has gone down.

2. What visualizations would help readers understand the trends? (5 points)

A scatter plot or a histogram would help people visualize and understand these trends. The scatterplot would be the best was to measure because I could use a line of best fit to accurately show if the response time has gotten worse or if it nearly an exaggeration. 

3. What additional context or data would you need to make this a complete story? (5 points)

Other context I would need to establish would be if the emergency response force has been losing volunteers or staff and try to establish a possible corelation between a dwindling staff and slower emergency response times.

### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Before submitting, clear your environment and run the entire notebook

Good luck!
