---
title: "HW14_USATodayIndex"
output: html_document
author: William Deffaa
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Overview

In this exercise, we will calculate the **USA TODAY Diversity Index** for Maryland cities with a population of at least 10,000 using 2020 Census race and ethnicity data. The Diversity Index reflects the probability that two randomly selected individuals from a city are of different races or ethnicities.

A score of 0 indicates no diversity (everyone is the same race/ethnicity), while a score closer to 100 indicates higher diversity. The **most diverse city** has the highest index, reflecting a balanced mix of several racial and ethnic groups. The **least diverse** city likely has one overwhelmingly dominant racial/ethnic group.

## Task 1: Load and Describe the Data (3 points)

Run the following code to load the data, then examine it and describe what it contains. Based on the raw numbers alone, which cities stand out as diverse to you?

The main cities that stood out to me in terms of diversity were Baltimore City, Towson, Germantown, and Ellicott City. 

```{r load-data}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/usat_diversity_index.csv")

# Preview structure
View(md_cities)
```

## Task 2: Calculate Race Percentages

```{r race-percentages}
md_cities <- md_cities |>
  mutate(
    White_pct = White / TotalPop,
    Black_pct = Black / TotalPop,
    Native_pct = Native / TotalPop,
    Asian_pct = Asian / TotalPop,
    NHPI_pct = NHPI / TotalPop
  )
```

## Task 3: Square the Race Percentages

We need to square the race percentages in order to calculate the probability that any two people chosen at random from a given census area are of different races or are Hispanic or not.

```{r square-race}
md_cities <- md_cities |>
  mutate(
    White_pct_sq = White_pct^2,
    Black_pct_sq = Black_pct^2,
    Native_pct_sq = Native_pct^2,
    Asian_pct_sq = Asian_pct^2,
    NHPI_pct_sq = NHPI_pct^2
  )
```


## Task 4: Calculate Race Probabilities (3 points)

The race probability is the sum the squared probabilities for each of the races. It represents the probability that two people randomly chosen from the same city will be of the same race. Which city has the highest race probability, and why do you think that is?

The city that has the highest race probability is Summerfield, which could be explained because it is just east of D.C. and contains the Washington Commanders stadium.

```{r race-prob}
md_cities <- md_cities |>
  mutate(
    Race_Prob = White_pct_sq + Black_pct_sq + Native_pct_sq + Asian_pct_sq + NHPI_pct_sq
  ) |> 
  arrange(desc(Race_Prob))

View(md_cities)
```


## Task 5: Calculate Ethnicity Percentages and Their Squares (3 points)

Hispanic origin is not a race but an ethnicity or nationality – and a separate Census question. The probability someone is Hispanic or not Hispanic must be calculated separately from race, and Hisp_Prob is the probability that two randomly selected people will be Hispanic or not. That's an important distinction, because unlike race probability, this measures the likelihood of a binary question: that both random people will be Hispanic. What do you notice about the cities with the *lowest* Hisp_Prob scores?

Most of the cities that have the lowest Hisp_Prob scores are cities that surround Washington D.C.

```{r ethnicity}
md_cities <- md_cities |>
  mutate(
    HispTotal = HISPANIC + NotHispanic,
    Hisp_pct = HISPANIC / HispTotal,
    NotHisp_pct = NotHispanic / HispTotal,
    Hisp_pct_sq = Hisp_pct^2,
    NotHisp_pct_sq = NotHisp_pct^2,
    Hisp_Prob = Hisp_pct_sq + NotHisp_pct_sq
  ) |> 
  arrange(Hisp_Prob)

View(md_cities)
```

## Task 6: Multiply Race and Ethnicity Probabilities

Here we multiply the two probabilities together to get the probability that any two random people have the SAME race and ethnicity.

```{r combined-prob}
md_cities <- md_cities |>
  mutate(
    RaceEth_Prob = Race_Prob * Hisp_Prob
  )
```


## Task 7: Calculate Diversity Index (3 points)

The final step is to calculate the probability that two people randomly chosen WILL NOT BE OF THE SAME RACE OR ETHNICITY. In other words, the highest index scores will be cities where those two random people will be from different backgrounds. Put simply, more diverse. Which cities are the top 3, and where are they located? Look at the bottom of the list - which city is the least diverse, and where is it?

The cities that are in the top 3 are Langley Park, Wheaton, and Adelphi. These cities are all located just around D.C. and are relatively close to College Park. The city that has the lowest diversity index is Lake Shore, which is located just outside Annapolis, near Gibson Island. This is a place were some of the wealthiest people in the state live or have beach/bay houses.

```{r diversity-index}
md_cities <- md_cities |>
  mutate(
    Diversity_Index = (1 - RaceEth_Prob) * 100
  ) |> 
  arrange(desc(Diversity_Index))

View(md_cities)
```

### Pitch a story (8 points)

Given the data that you've produced here, write a headline and lede of the most interesting story from this exercise, and add below that details on how you would report it out: who you'd talk to and the questions you'd ask. When you describe the Diversity Index values, do so in a similar way to how USA Today did; the raw values themselves won't mean much to readers.

Cities in Maryland Around D.C. are the most Diverse, While Annapolis and Surrounding Cities are the Least.

Diversity indexes cite cities around Washington D.C. as being the most diverse in Maryland, while cities around Annapolis are cited as the least diverse cities. The most diverse cities according the diversity index are Langley Park, Wheaton, and Adelphi, all cities around Washington D.C.. The least diverse city is Lake Shore, located just outside of Annapolis. The diversity index measures the probability that two people randomly chosen are not the same race or ethnicity, the index is from 0-100 and the higher the value, the more diverse and area is.
