---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

The most important categories are Homicide, Theft, and the various types of Assults that are listed. One of the categories that makes it hard to summarize is the Homicide category, as unlike many of the other broader categories, the Homicide category does not have subcategories like Assult or Robbery. 

### Task 4: Calculating Averages

Calculate the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_averages <- pg_crime %>%
  group_by(category) %>%
  mutate(total_incidents = rowSums(across(Jan:Dec))) %>%
  mutate(average_incidents = total_incidents / 12)
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

Accidents had the highest average number of incidents per moth with Auto Stolen in second. The lowest categories were Homicide and B & E Other. This shows that most of the crime are Accidents, Theft, and Stolen Auto.


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages %>% 
  group_by(category) %>% 
  mutate(june_change = (`Jun` - `average_incidents`)/`average_incidents` *100) %>%
  mutate(july_change = (`Jul` - `average_incidents`)/`average_incidents` *100) %>%
  mutate(august_change = (`Aug` - `average_incidents`)/`average_incidents` *100)
  

```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

Accidents lower in June and July, then go back up in August. Homicides see a major increase in June, July, and August and so do B&E Other as well. Sex Offense incidents are up in June and July (more so July), and then see a big drop in August. Violent crimes such as Homicides and Sex Offense incidents are much more prone to seeing rises in the summer due to more people being out and about, therefore leading to a higher probability of crime being committed.  

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

A big rise in violent crimes such as Homoicides and Sex Offenses has been shown in the summer months, while accidents see a drop as opposed to non-summer months.

Homicides have seen a rise in the summer months, with a 37.14% increase in June and a 20% increase in July and August. In addition, Sex Offenses also see a rise in some of the summer months, with June and july seeing rising numbers with increases of 14% and 22% respectively. However, August sees a large decline, with a drop of 18% from the average number of incidents per month. One interesting statistic is the decrease in Accidents in some of the summer months, with July seeing a 12% decrease than the average. This data is important because it connects the theory that violent crimes seem to have a upward trend come the warmer, summer months with more people being outside than usual, which leads to the larger possibility for crime. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
