---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv") %>% clean_names()
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities %>% mutate(violent_rate_2019 = violent2019/pop2019 * 100) %>% 
  mutate(property_rate_2019 = property2019/pop2019 * 100) %>% 
  mutate(violent_rate_2020 = violent2020/pop2020 * 100) %>% 
  mutate(property_rate_2020 = property2020/pop2020 * 100)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

Crime is much higher in densely populated cities (not much of a surprise) and in more rural areas it is a lot lower with some even recording 0 for property and violent crime rates. Surprisingly, Elkton and other Maryland cities such as Ocean city have much higher property crime rates in 2019 than Baltimore.


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates %>% 
  mutate(property_rate_2020 - property_rate_2019/property_rate_2019 * 100) %>% 
  mutate(violent_rate_2020 - violent_rate_2019/violent_rate_2019 * 100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

A major story of the data is showing a big decrease in both violent crime rates and property crime rates with many having decreases of -95% or more. 
(If I misread a step or messed up please let me know) 

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Many Maryland cities have seen a large drop in violent and property crime rates from 2019 to 2020, could this be a result of the COVID-19 pandemic or has this been a trend over the past 5-10 years. Even violent cities like Baltimore have seen a drop in violent crime rates as well.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10: Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile? Calculate the Interquartile Range of the 2020 violent crime rate using `summarize()` and R's `IQR()` function. The syntax is similar to how `mean()` works.

```{r}
md_cities_percentiles <- md_cities_rates %>% mutate(percentile = ntile(violent_rate_2020, 100)) %>%
summarize(iqr_value = IQR(violent_rate_2020))
```

11 cities are at or above the 90th percentile. The IQR value is 0.4198 or 41.98%.
