---
title: "HW7_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
```


### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
candy_mnm <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset
```{r}
sample_5 <- candy_mnm %>% sample_n(5)
```


### Task 4: Calculate the mean and standard deviation of the sample, using either the red or black number of M&Ms
```{r}
sample_5 %>% summarize(mean = mean(red))
  sample_5 %>% summarize(standardd = sd(red))
```


### Task 5: Calculate the mean and standard deviation of whole table, using the same measure as in Task 4
```{r}
candy_mnm %>% summarize(standardd = sd(red))
candy_mnm %>% summarize(mean = mean(red))
```


### Task 6: Compare the M&M statistics

Compare the two sets of means and standard deviations. Are they similar?

The standard deviation of the sample is somewhat similar to the standard deviation of the full data, however, the mean of the sample seems to be much lower than the mean of the overall data. 
### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

The biggest thing I noticed was that the mean was lower in the sample than the full data, this mainly tells me that while the standard deviation of a sample could possibly be accurate when also refering to a total dataset, measuring the mean via a sample will not get entirely accurate numbers that reflect the actual data.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
