---
title: "final_project"
output: html_document
---

Polished Project

(All narratives are below the code that was written)

```{r}
library(tidyverse)
```
```{r}
Movie_Reviews_FinalProj %>% summarize(standardd = sd(Adj_RT_Rating))
  Movie_Reviews_FinalProj %>% summarize(mean = mean(Adj_RT_Rating))
```
```{r}
Movie_Reviews_FinalProj %>%
  group_by(Year_Released) %>%
  summarize(avg_rating = mean(Adj_RT_Rating, na.rm = TRUE)) %>%
  ggplot(aes(x = Year_Released, y = avg_rating)) +
    geom_col(fill = "red", color = "black") +
    labs(title = "Average Rotten Tomato Rating by Year",
         x = "Year Released", y = "Average Rating")

```

```{r}
Movie_Reviews_FinalProj %>% summarize(standardd = sd(LTBX_Rating))
  Movie_Reviews_FinalProj %>% summarize(mean = mean(LTBX_Rating))
```

```{r}
Movie_Reviews_FinalProj %>%
  group_by(Year_Released) %>%
  summarize(avg_rating = mean(LTBX_Rating, na.rm = TRUE)) %>%
  ggplot(aes(x = Year_Released, y = avg_rating)) +
    geom_col(fill = "lightblue", color = "black") +
    labs(title = "Average LetterBoxed Rating by Year", x = "Year Released", y = "Average Rating")
```

First I wanted to present bar graphs of the averages of LetterBoxed and Rotten Tomatoes ratings of movies by year. My dataset contains movies from 2024 to 1934 and I tried to have every year be equally represented, trying to have 5 movies for each year. Through my created bar charts, we can see that average Rotten Tomatoes ratings drop around the 1980s mark, then starting to climb back up right before 2000, and finally after leveling off in the 2000s, they see a steady increase. With LetterBoxed ratings, the average ratings seem to fluctuate more in the earlier years prior to taking a dip after 1975 and then leveling out. Similar to Rotten Tomatoes average ratings, LetterBoxed average ratings seem to steadily increase during the 2000s. 

```{r}
Movie_Reviews_FinalProj %>% summarize(correlation = cor(Adj_RT_Rating, Year_Released, method = "pearson"))
```

```{r}
Movie_Reviews_FinalProj %>%
  ggplot() +
  geom_point(aes(x = Year_Released, y = Adj_RT_Rating)) +
  geom_smooth(aes(x = Year_Released, y = Adj_RT_Rating), method = "lm") +
  labs(title = "Correlation Between Rotten Tomatoes Ratings and Year Released", x = "Year Released", y = "Rotten Tomatoes Ratings")
```

```{r}
Movie_Reviews_FinalProj %>% summarize(correlation = cor(LTBX_Rating, Year_Released, method = "pearson"))
```

```{r}
Movie_Reviews_FinalProj %>%
  ggplot() +
  geom_point(aes(x = Year_Released, y = LTBX_Rating)) +
  geom_smooth(aes(x = Year_Released, y = LTBX_Rating), method = "lm") + 
  labs(title = "Correlation Between LetterBoxed Ratings and Year Released", x = "Year Released", "LetterBoxed Ratings")
```

Next, I wanted to see the correlations between the two different movie reviewers I am using and the year in which the movies are released. For this test, I calculated the correlations between both rating systems I'm using and the year movies were released and then put the data for each ratings on a plot point chart, with a line of best fit on the graphs as well. For both of these correlation tests, they both came back with a negative correlation between ratings and movies by the year they were released, with Rotten Tomatoes showing a stronger negative correlation than LetterBoxed ratings. This indicates that throughout time, movies have received less higher ratings. 


```{r}
cor.test(Movie_Reviews_FinalProj$Adj_RT_Rating, Movie_Reviews_FinalProj$LTBX_Rating, method = "pearson")
```

```{r}
Movie_Reviews_FinalProj %>%
  ggplot() +
  geom_point(aes(x = LTBX_Rating, y = Adj_RT_Rating)) +
  geom_smooth(aes(x = LTBX_Rating, y = Adj_RT_Rating), method = "lm") +
  labs(title = "Correlation Between Rotten Tomatoes and LetterBoxed Ratings", x = "LetterBoxed Ratings", y = "Rotten Tomatoes Ratings")
```

For my first main test, I decided to test the correlation between Rotten Tomatoes ratings and LetterBoxed ratings. My hypothesis was that critics opinions, which are put into the data set via Rotten Tomatoes ratings, are often indicative of public opion, which is represented by LetterBoxed ratings in the data set. I used a correlation test in which I used the "cor.test" function to test the correlation between the two rating systems. When using this function, the correlation coefficient came out to 0.667, suggesting a strong positive linear relationship between the two, meaning they are aligned. Additionally, the p-value is shown to be lower than 0.05, in which we can reject the null hypothesis and say that there is a statistically significant correlation between Rotten Tomatoes and LetterBoxed ratings. In conclusion, my first hypothesis was correct, movie critics opinion is often indicative of public opinion.

```{r}
Movie_Reviews_FinalProj %>%
  ggplot(aes(x = Adj_RT_Rating, y = LTBX_Rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatterplot of Rotten Tomatoes vs LetterBoxed Ratings", x = "Rotten Tomatoes Ratings", y = "LetterBoxed Ratings")
```

```{r}
model <- lm(LTBX_Rating ~ Adj_RT_Rating, data = Movie_Reviews_FinalProj)

summary(model)

```

In addition to testing the correlation between Rotten Tomatoes and LetterBoxed ratings, I decided to test the linear regression and see how much Rotten Tomatoes scores influence LetterBoxed scores (public opinion). From my test, I was able to see a strong linear relationship between the two rating systems. According to the model, with every 1-point increase in a Rotten Tomatoes score (using the adjusted Rotten Tomatoes range), a LetterBoxed score increases nearly 0.35 points. According to the F-statistic, the model as a whole is highly significant.

```{r}
Movie_Reviews_FinalProj %>% 
  ggplot(aes(x = Year_Released, y = Adj_RT_Rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Scatterplot of Year of Movies Released vs Rotten Tomatoes Ratings", x = "Year Released", y = "Rotten Tomatoes Ratings")
```

```{r}
model <- lm(Adj_RT_Rating ~ Year_Released, data = Movie_Reviews_FinalProj)

summary(model)
```

```{r}
Movie_Reviews_FinalProj %>%
  ggplot(aes(x = Year_Released, y = LTBX_Rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +
  labs(title = "Scatterplot of Year of Movies Released vs Letterboxd Ratings", x = "Year Released", y = "Letterboxd Ratings") 
```

```{r}
model <- lm(LTBX_Rating ~ Year_Released, data = Movie_Reviews_FinalProj)

summary(model)
```

For my second major test, I wanted to test my hypothesis that critics (Rotten Tomatoes Ratings), are more favorable to older movies compared to newer movies. And on the other end, public opinion (Letterboxd Ratings), are more favorable to newer movies rather than old ones. I determined "older" movies to be pre 1975, as that year is one of the years marked on the earlier bar charts. To test this hypothesis, I took both Rotten Tomatoes and Letterboxd ratings and put them in a linear regression model with the Year_Released and attempted to see if either ratings have decreased or increased throughout time and if there has been a trend since 1975. For Rotten Tomatoes ratings (critics), they tend to favor older movies more, as newer movies have seen a slightly lower rating over time. However, the effect is still small and there is only a 11.4% variance, as shown by the Adjusted R-squared, suggesting other factors may contribute to this. When doing this same test with Letterboxd ratings (public opinion), while there is the same decline in ratings given by users for newer movies rather than older movies, only about 8% of the variation in these ratings are due to the year released, suggesting that public opinion on movies are not really determined by how old or new they are. In conclusion, while my hypothesis was correcy, with both of these models showing that both critics (Rotten Tomatoes) and overall public opinion (Letterboxd) seem to favor older movies due to the fact that they are higher rated than newer movies, it's more than likely that other external factors determine these ratings rather than simply people's opinions/bias toward older movies.

Story Pitch:

I would pitch a story about how critics and possible public opinion overall has been trending downward on newer movies. In addition, I would suggest possibly diving into how much critics (Rotten Tomatoes) truly contribute to getting people out there to see "good" movies, possibly contributing to thier success amongst the general public. My analysis and findings would be that while public and critical opinion has been trending downward on newer movies, there are many more factors that contribute to this phenomenon rather than just a bias toward older movies. Additionally, I would also talk about how critics opinions often align very closely to overall public opinion on a film, debunking the theory that critics may be pretentious and rate movies more harshly than the overall public. Unfortunately, some problems that me be of concern would be the fact that the lower ratings of movies over time wouldn’t be entirely due to bias/preference people have toward older movies, but due to many other external factors and these factors would be difficult to find. For a reporting plan I would try to first of all expand on my list of movies, as while I think this was a good list to start out with, it would need to be bigger and the movies generated would need to be more methodical. Secondly, I would try and see what other factors I would look into to further understand why ratings for newer movies are down both by critics and the general public. And finally, I would try to look at other public movie rating sites to see if the same correlation would exist just like my original test did which would help me get a futher understanding of public opinion rather than just getting it from one site.  
